\TESS Photometric Mapping of a Terrestrial Planet in the Habitable Zone: Detection of Clouds, Oceans, and Continents

### 2019-03-28 tex/ms.tex
intro

Of the many challenges and opportunities facing the field of exoplanets in the future, the prospect of mapping the surface of a potentially habitable planet is one of the most exciting -- and one of the most difficult. 
Present-day missions including the Transiting Exoplanet Survey Satellite \citep[\TESS; ][]{Ricker2015} aim to discover small terrestrial planets around bright, nearby stars. 
Once these targets are found, much more detailed observational characterization must follow to evaluate their suitability for harboring life. 
The discovery of surface features including continents and oceans would be a critical hint towards the habitability of an alien world.

The most straightforward method of observing features on a terrestrial exoplanet would be to directly image the planet at sufficient resolution. 
However, such an observation would require an effective telescope aperture orders of magnitude larger than currently available as well as extremely advanced coronagraphic optics, making it a far-future goal.

Surface mapping can also be done by observing rotational phase-dependent variations of the planetary signal in precise photometric time series. 
This is a complex data analysis problem: strong degeneracies are involved when transforming a 1-dimensional time series into a 2-dimensional surface map. 
Interpretation is also challenging in the face of many potential explanations for a poorly resolved observed feature.

\todo{(Summary of previous mapping work:)}\citep{Knutson2007,Majeau2012,deWit2012}

@@

Our full model for each light curve is simply the product of the 
systematics model $\mathbf{p}_n = \mathbf{B}\mathbf{w}_n$ (different for each target) and
the \starry model $\mathbf{s} = \mathbf{A}\mathbf{y}$ (shared by all targets). Since we happen
to know the exact distance $\mathbf{r}$ between \TESS and Sol d at all times
(\S\ref{sec:orbit}), we also include the inverse square of $\mathbf{r}$ 
as a multiplicative term. The model for the flux timeseries
in the $n^\mathrm{th}$ postage stamp is therefore
%
\begin{align}
    \label{eq:model}
    \mathbf{m}_n = (\mathbf{r}^{-2}) \circ (\mathbf{B} \mathbf{w}_n) \circ (\mathbf{A} \mathbf{y})
\end{align}
%
where $\circ$ denotes the element-wise product of two vectors.

@@

The problem of inferring a two-dimensional map from a one-dimensional
timeseries is famously ill-posed given its large null space and many complex
degeneracies \citep[e.g.,][]{CowanFuentesHaggard2013}. In fact, for a 
uniformly illuminated body, its rotational light curve encodes at most
$2l_\mathrm{max}$ modes (one sine term and one cosine term per frequency), 
where $l_\mathrm{max}$ is the highest spherical
harmonic degree of features on the surface of the body. The number of modes
\emph{on the surface of the body}, however, increases quadratically 
as $(l_\mathrm{max} + 1)^2$. This means that for a body whose surface is
perfectly described by a spherical harmonic map of degree $l_\mathrm{max} = 10$,
which is described by 121 coefficients, only 20 linear combinations
of those coefficients are actually projected onto the light curve. The
problem is fundamentally non-invertible, since most of the information
one wishes to extract never makes it into the light curve in the first place.

>>>>

Of the many challenges and opportunities facing the field of exoplanets in the future, 
the prospect of mapping the surface of a potentially habitable planet is one of 
the most exciting -- and one of the most difficult. 
Present-day missions including the Transiting Exoplanet Survey Satellite 
\citep[\TESS; ][]{Ricker2015} aim to discover small terrestrial planets around 
bright, nearby stars. 
Once these targets are found, more detailed observational characterization 
must follow to evaluate their suitability for harboring life. 
The discovery of surface features such as continents and oceans would be a 
critical step toward understanding the habitability of an alien world.

The most straightforward method of observing features on a terrestrial exoplanet 
would be to directly image the planet at sufficient resolution. 
However, such an observation would require an effective telescope aperture many orders 
of magnitude larger and equipped with far more advanced coronagraphic optics
than what we have available today.

Fortunately, surface mapping can also be done indirectly by observing rotational 
phase-dependent variations of the planetary signal in precise photometric time series. 
This is a complex data analysis problem: strong degeneracies are involved when 
transforming a 1-dimensional time series into a 2-dimensional surface map
\citep{CowanFuentesHaggard2013}. 
Interpretation is also challenging in the face of many potential explanations 
for a poorly resolved surface feature. Nevertheless, this technique has
already proven successful in the mapping of large dipolar features on the surfaces
of hot giant planets, most notably the hot Jupiter HD189733b
\citep{Knutson2007,Majeau2012,deWit2012}. More recently, this technique has also
been applied to the potentially terrestrial planet 55 Cancri e, suggesting
a large longitudinal offset in its peak surface emission due to either
lava flows on the surface or strong atmospheric recirculation
\citep{Demory2016,Demory2016b,Hammond2017}.

@@

Our full model for each light curve is simply the product of the 
systematics model $\mathbf{p}_n = \mathbf{B}\mathbf{w}_n$ (different for each target) and
the \starry model $\mathbf{s} = \mathbf{A}\mathbf{y}$ (shared by all targets). Since we happen
to know the exact distance $\mathbf{r}$ between \TESS and Sol d at all times
(\S\ref{sec:orbit}), we also include the inverse square of $\mathbf{r}$ 
as a multiplicative term. The model for the flux time series
in the $n^\mathrm{th}$ postage stamp is therefore
%
\begin{align}
    \label{eq:model}
    \mathbf{m}_n = (\mathbf{r}^{-2}) \circ (\mathbf{B} \mathbf{w}_n) \circ (\mathbf{A} \mathbf{y})
\end{align}
%
where $\circ$ denotes the element-wise product of two vectors.

@@

The problem of inferring a two-dimensional map from a one-dimensional
time series is famously ill-posed given its large null space and many complex
degeneracies \citep[e.g.,][]{CowanFuentesHaggard2013}. In fact, for a 
uniformly illuminated body, its rotational light curve encodes at most
$2l_\mathrm{max}$ modes (one sine term and one cosine term per frequency), 
where $l_\mathrm{max}$ is the highest spherical
harmonic degree of features on the surface of the body. The number of modes
\emph{on the surface of the body}, however, increases quadratically 
as $(l_\mathrm{max} + 1)^2$. This means that for a body whose surface is
perfectly described by a spherical harmonic map of degree $l_\mathrm{max} = 10$,
which is described by 121 coefficients, only 20 linear combinations
of those coefficients are actually projected onto the light curve. The
problem is fundamentally non-invertible, since most of the information
one wishes to extract never makes it into the light curve in the first place.
